Timeliness of security updates for Android

### 2014-08-11 security_scoring.tex
Consistently use the term 'operator' rather than 'carrier'

The scores across the whole of android are that \daMeanInsecurityPerc\ of devices are exposed to known root exploits.
There are on average \daMeanOutstandingVulnerabilities\ outstanding vulnerabilities not fixed on any device.
Only on average \daUpdatednessPerc\ of devices run the most recent version of Android.
This gives a security score of \daSecurityScore / 10.
\daTabSecScoressummary
However there are a wide variety of scores depending on the source of the device.
There have been many reports that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices.
Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices.
\daTabSecScoresmanufacturer
Different manufacturers have very different scores, Table~\ref{tab:sec_manufacturer} shows the scores for the \daNumSigManufacturers\ manufacturers with a significant presence in our data with \daSecScoreBestmanufacturer\ (\daSecScoreBestmanufacturerScore\ / 10) doing best and \daSecScoreWorstmanufacturer\ (\daSecScoreWorstmanufacturerScore\ / 10) doing worst.
Manufacturers are considered significant if we have data from at least \daSigNumDevices\ devices and at least \daSigNumDeviceDays\ days of contributions.
Additionally for $m$ and $u$ we ignore days with less than \daSigNumDevicesDay\ devices contributing to that day's score.
\daTabSecScoresmodel
Even within manufacturers different models can have very different update behaviours and hence security.
Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric with \daSecScoreBestmodel\ (\daSecScoreBestmodelScore\ / 10) doing best and \daSecScoreWorstmodel\ (\daSecScoreWorstmodelScore\ / 10) doing worst.

>>>>

The scores across the whole of android are that \daMeanInsecurityPerc\ of devices are exposed to known root exploits.
There are on average \daMeanOutstandingVulnerabilities\ outstanding vulnerabilities not fixed on any device.
Only on average \daUpdatednessPerc\ of devices run the most recent version of Android.
This gives a security score of \daSecurityScore / 10.
\daTabSecScoressummary
However there are a wide variety of scores depending on the source of the device.
There have been many reports that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices.
Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices.
\daTabSecScoresmanufacturer
Different manufacturers have very different scores, Table~\ref{tab:sec_manufacturer} shows the scores for the \daNumSigManufacturers\ manufacturers with a significant presence in our data with \daSecScoreBestmanufacturer\ (\daSecScoreBestmanufacturerScore\ / 10) doing best and \daSecScoreWorstmanufacturer\ (\daSecScoreWorstmanufacturerScore\ / 10) doing worst.
Manufacturers are considered significant if we have data from at least \daSigNumDevices\ devices and at least \daSigNumDeviceDays\ days of contributions.
Additionally for $m$ and $u$ we ignore days with less than \daSigNumDevicesDay\ devices contributing to that day's score.

\daTabSecScoresmodel
Even within manufacturers different models can have very different update behaviours and hence security.
Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric with \daSecScoreBestmodel\ (\daSecScoreBestmodelScore\ / 10) doing best and \daSecScoreWorstmodel\ (\daSecScoreWorstmodelScore\ / 10) doing worst.

\daTabSecScoresoperator
We also analysed different operators, for the \daNumSigOperators\ operators with a significant presence in our data Table~\ref{tab:sec_operator} shows the results \daSecScoreBestoperator\ (\daSecScoreBestoperatorScore\ / 10) doing best and \daSecScoreWorstoperator\ (\daSecScoreWorstoperatorScore\ / 10) doing worst.

### 2014-08-11 securityupdates.tex
Consistently use the term 'operator' rather than 'carrier'

\subsection{Android update process}
\begin{figure}
 \centering
 \def\svgwidth{\columnwidth}
 \import{figures/}{update_ecosystem.pdf_tex}
 \caption{Flow of updates between participants in the Android ecosystem.
 Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, numbers in brackets represent number of such entities in our data.}
 \label{fig:update_ecosystem}
\end{figure}
A simple update process would be for a developer, who wrote the program, to write an update and ship it to the user who installs it.
The process for Android is rather more complicated and is shown in Figure~\ref{fig:update_ecosystem}.
Android builds on various open source projects such as the Linux kernel, the OpenSSL and BouncyCastle cryptography libraries and various Apache Java libraries.
It also incorporates various drivers for different bits of hardware.
The Android platform is then built on top of those with kernel and userspace components by Google and the Open Handset Alliance. %TODO work out the exact arrangements.
Usually the code for each update is kept secret until after the update has been published.
Then the Manufacturers take stock Android and customise it with different themes, apps, drivers and so on and ship updates on, to the carrier.
The Carrier may then make further customisations and do further testing before shipping the update to the User.
Sometimes Manufactures ship direct to the user, sometimes the Manufacturer and Google are the same entity (at least for the purpose of a particular phone, such as with Nexus phones).
An analysis by Vidas et al.~\cite{Vidas2011} of the Android 2.1 to 2.2 update found that it took 11 months from when Google released 2.2 for the last device which they were investigating to get the update.
This multi-stage multi-party release process increases the exposure of Android devices to vulnerabilities as it takes a long time for them to be patched.

>>>>

\subsection{Android update process}
\begin{figure}
 \centering
 \def\svgwidth{\columnwidth}
 \import{figures/}{update_ecosystem.pdf_tex}
 \caption{Flow of updates between participants in the Android ecosystem.
 Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, numbers in brackets represent number of such entities in our data.}
 \label{fig:update_ecosystem}
\end{figure}
A simple update process would be for a developer, who wrote the program, to write an update and ship it to the user who installs it.
The process for Android is rather more complicated and is shown in Figure~\ref{fig:update_ecosystem}.
Android builds on various open source projects such as the Linux kernel, the OpenSSL and BouncyCastle cryptography libraries and various Apache Java libraries.
It also incorporates various drivers for different bits of hardware.
The Android platform is then built on top of those with kernel and userspace components by Google and the Open Handset Alliance. %TODO work out the exact arrangements.
Usually the code for each update is kept secret until after the update has been published.
Then the Manufacturers take stock Android and customise it with different themes, apps, drivers and so on and ship updates on, to the operator.
The Operator may then make further customisations and do further testing before shipping the update to the User.
Sometimes Manufactures ship direct to the user, sometimes the Manufacturer and Google are the same entity (at least for the purpose of a particular phone, such as with Nexus phones).
An analysis by Vidas et al.~\cite{Vidas2011} of the Android 2.1 to 2.2 update found that it took 11 months from when Google released 2.2 for the last device which they were investigating to get the update.
This multi-stage multi-party release process increases the exposure of Android devices to vulnerabilities as it takes a long time for them to be patched.
